---
title: "overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bnRep)
```

The R package `bnRep` includes the largest repository of Bayesian networks, which were all collected from recent academic literature in a variety of fields! If you are using any Bayesian network from `bnRep` you should cite:

Leonelli, M (2024). "bnRep: A repository of Bayesian networks from the academic literature." ArXiv 24....

```
@Article{,
    title = {bn{R}ep: A repository of {B}ayesian networks from the academic literature},
    author = {Manuele Leonelli},
    journal = {Arxiv},
    year = {2024}
  }
```

Go to (link here) to explore the repository online!

## Contribution

If you are interested in having your Bayesian network included in `bnRep` you must prepare three objects:

 - the Bayesian network as a `bn.fit` object (if not created with `bnlearn` you can always use import functions, such as `read.bif()`);
 
 - an R file with the same name of the `bn.fit` object reporting the documentation of the Bayesian network;
 
 - a vector/excel file with the required details to include in the `bnRep_summary` object.
 
You can submit the required objects directly via github (e.g fork/pull), or via email.
 
If you struggle with any of these steps, please get in touch and I will try to help!

## Overview

`bnRep` includes over 200 Bayesian networks from more than 150 academic publications. It includes discrete, Gaussian and conditional linear Gaussian Bayesian networks, all stored as appropriate `bn.fit` objects from `bnlearn`. They can be exported for use to other software (e.g. Phython libraries) using functions from `bnlearn` such as `write.bif()`. Recall that in order to plot the associated DAG, one must first convert it to a graph object with `bn.net()` from the `bnlearn` package.

### Installation

```{r, eval = FALSE}
# Install stable version from CRAN:
install.packages("bnRep")

# Or the development version from GitHub:
remotes::install_github("manueleleonelli/bnRep")
```

### Usage

We will use the `lawschool` Bayesian network as an example. To load it in the environment simply call `data(lawschool)` and to then plot it (for instance using `graphviz.plot` from the `bnlearn` package)

```{r message=F, out.width="50%"}
library(bnRep)
library(bnlearn)
data("lawschool")
qgraph::qgraph(bn.net(lawschool))
```

Notice that the function `bn.net` function must be used in order to plot the network.

### Exploring bnRep

`bnRep` includes two features to explore the Bayesian networks in the repository:

 - `bnRep_summary`: a dataframe with important details about each network in the repository.
 
 - `bnRep_app`: a Shiny app to interactively explore `bnRep_summary` and filter the networks according to various criteria. The app is also available online at (link here).
 
Here's the columns of `bnRep_summary`:

```{r echo=F}
colnames(bnRep_summary)
```

### An overview of the Bayesian networks in bnRep

The following plots show some summary statistics of the repository.

```{r echo=F,message= F, out.width="50%"}
library(ggplot2)
library(dplyr)
library(stringr)
library(scales)
library(RColorBrewer)

# Assuming bnRep_summary is your data frame and Type is a factor column
# Creating the barplot with percentages on the Y-axis and labels on the bars
bnRep_summary %>%
  count(Type) %>%
  mutate(perc = n / sum(n) * 100) %>%
  ggplot(aes(x = Type, y = perc, fill = Type)) +
  geom_bar(stat = "identity", width = 0.7, color = "black", show.legend = FALSE) +
  geom_text(aes(label = paste0(round(perc, 1), "%")), vjust = -0.5, size = 5, color = "black") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +
  labs(title = "Bayesian networks by type",
       x = "Type",
       y = "Percentage") +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank()
  ) +
  scale_fill_brewer(palette = "Pastel1")
```

```{r echo=F,message= F, out.width="50%"}
bnRep_summary %>%
  count(Structure) %>%
  mutate(perc = n / sum(n) * 100) %>%
  ggplot(aes(x = Structure, y = perc, fill = Structure)) +
  geom_bar(stat = "identity", width = 0.7, color = "black", show.legend = FALSE) +
  geom_text(aes(label = paste0(round(perc, 1), "%")), vjust = -0.5, size = 5, color = "black") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 50)) +
  labs(title = "Bayesian networks by structure definition",
       x = "Type",
       y = "Percentage") +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank()
  ) +
  scale_fill_brewer(palette = "Pastel1")

```
```{r echo=F,message= F, out.width="50%"}
bnRep_summary %>%
  count(Probabilities) %>%
  mutate(perc = n / sum(n) * 100) %>%
  ggplot(aes(x = Probabilities, y = perc, fill = Probabilities)) +
  geom_bar(stat = "identity", width = 0.7, color = "black", show.legend = FALSE) +
  geom_text(aes(label = paste0(round(perc, 1), "%")), vjust = -0.5, size = 5, color = "black") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 60)) +
  labs(title = "Bayesian networks by probabilities definition",
       x = "Type",
       y = "Percentage") +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank()
  ) +
  scale_fill_brewer(palette = "Pastel1")

```

```{r echo=F,message= F, out.width="50%"}
bnRep_summary %>%
  count(Graph) %>%
  mutate(perc = n / sum(n) * 100) %>%
  ggplot(aes(x = Graph, y = perc, fill = Graph)) +
  geom_bar(stat = "identity", width = 0.7, color = "black", show.legend = FALSE) +
  geom_text(aes(label = paste0(round(perc, 1), "%")), vjust = -0.5, size = 5, color = "black") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 70)) +
  labs(title = "Bayesian networks by graph type",
       x = "Type",
       y = "Percentage") +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank()
  ) +
  scale_fill_brewer(palette = "Pastel1")

```

```{r echo=F,message= F, out.width="50%"}
custom_colors <- colorRampPalette(RColorBrewer::brewer.pal(9, "Pastel1"))(length(unique(bnRep_summary$Area)))


bnRep_summary %>%
  count(Area) %>%
  mutate(perc = n / sum(n) * 100) %>%
  ggplot(aes(x = Area, y = perc, fill = Area)) +
  geom_bar(stat = "identity", width = 0.7, color = "black", show.legend = FALSE) +
  geom_text(aes(label = paste0(round(perc, 1), "%")), vjust = -0.5, size = 3.5, color = "black") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 25)) +
  labs(title = "Bayesian networks by academic area",
       x = "Type",
       y = "Percentage") +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 60, hjust = 1),
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank()
  ) +
  scale_fill_manual(values = custom_colors) 

```


```{r echo=F,message= F, out.width="50%"}
bnRep_summar <- bnRep_summary
bnRep_summar <- bnRep_summar %>%
  mutate(Year = ifelse(Year <= 2019, "2019 and earlier", as.character(Year)))

# Convert Year into a factor for plotting
bnRep_summar$Year <- factor(bnRep_summar$Year, levels = c("2019 and earlier", sort(unique(bnRep_summar$Year[bnRep_summar$Year != "2019 and earlier"]))))


bnRep_summar %>%
  count(Year) %>%
  mutate(perc = n / sum(n) * 100) %>%
  ggplot(aes(x = Year, y = perc, fill = Year)) +
  geom_bar(stat = "identity", width = 0.7, color = "black", show.legend = FALSE) +
  geom_text(aes(label = paste0(round(perc, 1), "%")), vjust = -0.5, size = 5, color = "black") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 35)) +
  labs(title = "Bayesian networks by year",
       x = "Type",
       y = "Percentage") +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank()
  ) +
  scale_fill_brewer(palette = "Pastel1")

```

```{r echo=F,message= F, out.width="50%"}
unique_journals_df <- bnRep_summary %>%
  distinct(Reference, .keep_all = TRUE) %>%
  group_by(Journal) %>%
  filter(n() >= 3) %>%
  ungroup()

unique_journals_df <- unique_journals_df %>%
  mutate(Journal = stringr::str_replace_all(Journal, "\\s", "\n")) 

# Create a barplot with counts instead of percentages
unique_journals_df %>%
  count(Journal) %>%
  ggplot(aes(x = Journal, y = n, fill = Journal)) +  # y = n for counts
  geom_bar(stat = "identity", width = 0.7, color = "black", show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.5, size = 5, color = "black") +  # Show counts on top of bars
  labs(title = "Bayesian Networks by Journal (having at least 3)",
       x = "Journal",
       y = "Count") +
  theme_minimal(base_size = 15) +
theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 0, hjust = 0.5),  # No rotation, centered
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank()
  ) +
  scale_fill_brewer(palette = "Pastel1")+
  scale_y_continuous(limits = c(0, 7))


```


```{r echo = F, out.width="50%"}


# Create the histogram with log10 scale on the x-axis
ggplot(bnRep_summary, aes(x = Nodes)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black", alpha = 0.7) +  # Customize the appearance
  scale_x_log10() +  # Set x-axis to log10 scale
  labs(title = "Distribution of Nodes (Log10 Scale)",
       x = "Number of Nodes (log10 scale)",
       y = "Count") +
  theme_minimal(base_size = 15) +  # Use a clean theme with larger base text size
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),  # Center the title
    axis.title.x = element_text(face = "bold"),  # Make x-axis title bold
    axis.title.y = element_text(face = "bold"),  # Make y-axis title bold
    panel.grid.major = element_line(color = "gray80"),  # Customize grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines for a cleaner look
    axis.text.x = element_text(angle = 45, hjust = 1)  # Angle the x-axis labels for better readability
  ) +
  scale_fill_brewer(palette = "Set2")  # Use a custom color palette


```

```{r echo =F, out.width="50%"}
ggplot(bnRep_summary, aes(x = Arcs)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black", alpha = 0.7) +  # Customize the appearance
  scale_x_log10() +  # Set x-axis to log10 scale
  labs(title = "Distribution of Arcs (Log10 Scale)",
       x = "Number of Arcs (log10 scale)",
       y = "Count") +
  theme_minimal(base_size = 15) +  # Use a clean theme with larger base text size
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),  # Center the title
    axis.title.x = element_text(face = "bold"),  # Make x-axis title bold
    axis.title.y = element_text(face = "bold"),  # Make y-axis title bold
    panel.grid.major = element_line(color = "gray80"),  # Customize grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines for a cleaner look
    axis.text.x = element_text(angle = 45, hjust = 1)  # Angle the x-axis labels for better readability
  ) +
  scale_fill_brewer(palette = "Set2") 
```
